Software for the analysis and visualization of deep mutational scanning data

### 2015-03-24 paper/BMC_Bioinformatics_2015/ResponseToReviewers.tex

3. Given the substantial increase in runtime and computational complexity of the Bayesian/MCMC approach relative to the simple ratio approach (hours vs fraction of a second), it might be helpful to provide a bit more intuition about what is gained. Is it simply a ``regularization'' where high ratios obtained from low counts are pushed back towards the prior? If so, does the framework really provide more than what could be obtained with a much simpler regularization heuristic?

\begin{response}
The reviewer asks the eminently reasonable question, ``Is the computational complexity of the Bayesian inference worth it?'' The revised manuscript includes a new subsection entitled \textsl{Is the Bayesian inference worthwhile?} directly after the description of the preference inference on simulated data. The honest answer to this question is equivocal -- the Bayesian inference appears to always perform as well or better than simple ratio calculations, but the improvements are non-negligible only in some situations. Hopefully the discussion in this new subsection provides a fair discussion of this topic.

Note that even when ratio estimation (which is also implemented in \texttt{dms\_tools} via the \texttt{--ratio\_estimation} option) is just as good, the visualization capabilities of \texttt{dms\_tools} will still be useful.
\end{response}

4. It’s not clear whether \texttt{dms\_inferprefs} outputs the probability of $\phi_{r,x} < 1$ (or $\phi_{r,x} \ne 1$). This is likely to be a useful feature.

\begin{response}
The \texttt{dms\_inferprefs} program does not output the posterior probability of $\phi_{r,x} \ne 1$ (or $< 1$), but it does output the 95\% credible interval for each $\pi_{r,x}$ value. The reason for not outputting the posterior probabilities of $\phi_{r,x} \ne 1$ is that the hypothesis that all amino acids have equal preferences at a given site does not represent a realistic view of proteins, as it is doubtful that there is any site in any protein where all 20 amino acids are equally preferred. Therefore, rather than providing a posterior probability that such an unrealistic hypothesis is true, it is more informative to simply output the credible interval of the preference for each amino acid at each site. This situation differs for the differential preferences discussed in the reviewer's next point -- for the differential preferences, it \emph{is} a plausible hypothesis that most preferences will be equal under two selection pressures, so computing posterior probabilities about this hypothesis is potentially useful.
\end{response}

@@

5. It's not clear whether the probability values for $\Delta\pi_{r,x} > 0$ or $< 0$ for \texttt{dms\_inferdiffprefs} should be interpreted as nominal or corrected. Providing false discovery rate estimates might be helpful for interpretation (also applies to 4.).

\begin{response}
The posterior probabilities $\Pr\left(\Delta\pi_{r,x} > 0 \mid \mbox{data}\right)$ reported by \texttt{dms\_inferdiffprefs} are \emph{not} corrected. However, note that these are Bayesian posterior probabilities, \emph{not} classical $P$-values. By default, \texttt{dms\_inferdiffprefs} places a prior over the $\Delta\pi_{r,x}$ values that biases them towards zero (see Equation 31 of the manuscript). Therefore, it is not clear that these posterior probabilities need to be corrected -- there are arguments for why Bayesian posterior probabilities inferred with appropriate priors do not need to be corrected (see \url{http://www.stat.columbia.edu/~gelman/research/published/multiple2f.pdf}). Rather than discuss these complex and incompletely resolved statistical questions, we have simply added the following sentence:
\begin{quote}
In addition, \texttt{dms\_inferdiffprefs} also creates text files that give the posterior probability that $\Delta\pi_{r,x} > 0$ or $< 0$ (these posterior probabilities are \emph{not} corrected to account for the fact that multiple sites are typically being examined, although be default the inferences are made using the regularizing prior in Equation 31).
\end{quote}
\end{response}

### 2015-03-24 paper/BMC_Bioinformatics_2015/ResponseToReviewers.tex

5. It's not clear whether the probability values for $\Delta\pi_{r,x} > 0$ or $< 0$ for \texttt{dms\_inferdiffprefs} should be interpreted as nominal or corrected. Providing false discovery rate estimates might be helpful for interpretation (also applies to 4.).

\begin{response}
The posterior probabilities $\Pr\left(\Delta\pi_{r,x} > 0 \mid \mbox{data}\right)$ reported by \texttt{dms\_inferdiffprefs} are \emph{not} corrected. However, note that these are Bayesian posterior probabilities, \emph{not} classical $P$-values. By default, \texttt{dms\_inferdiffprefs} places a prior over the $\Delta\pi_{r,x}$ values that biases them towards zero (see Equation 31 of the manuscript). Therefore, it is not clear that these posterior probabilities need to be corrected -- there are arguments for why Bayesian posterior probabilities inferred with appropriate priors do not need to be corrected (see \url{http://www.stat.columbia.edu/~gelman/research/published/multiple2f.pdf}). Rather than discuss these complex and incompletely resolved statistical questions, we have simply added the following sentence:
\begin{quote}
In addition, \texttt{dms\_inferdiffprefs} also creates text files that give the posterior probability that $\Delta\pi_{r,x} > 0$ or $< 0$ (these posterior probabilities are \emph{not} corrected to account for the fact that multiple sites are typically being examined, although be default the inferences are made using the regularizing prior in Equation 31).
\end{quote}
\end{response}

6. The function of \texttt{dms\_merge} isn't explained in the paper. Is it simply averaging the two replicate profiles? Could additional statistical power be gained by explicitly modeling replicates within the Bayesian framework?

\begin{response}
The revised manuscript now briefly explains \texttt{dms\_merge} and \texttt{dms\_correlate}. To jointly analyze replicates of the same experiment within the Bayesian framework, the counts from the replicates can simply be summed before inputting them to \texttt{dms\_inferprefs}. However, it may sometimes be desirable to analyze replicates individually with \texttt{dms\_inferprefs} and then combine and compare them using \texttt{dms\_merge} and \texttt{dms\_correlate}. The reason is that some of the replicate-to-replicate variability may arise from experimental factors distinct from statistical noise -- such factors are not easily incorporated into the Bayesian framework, since they are due to experimental rather than statistical sources.
\end{response}

### 2015-03-24 paper/BMC_Bioinformatics_2015/bmc_bioinformatics_article.tex
Responded to reviewer 1 point 6

\subsubsection*{Implementation.}
The program \texttt{dms\_inferprefs} in the \texttt{dms\_tools} package infers the preferences by using \texttt{pystan}~\cite{pystan-software:2014} to perform MCMC over the posterior defined by the product of the likelihoods and priors in Equations \ref{eq:pr_nrerrpre}, \ref{eq:pr_nrerrpost}, \ref{eq:pr_nrpre}, \ref{eq:pr_nrpost}, \ref{eq:pr_pir}, \ref{eq:pr_mur}, \ref{eq:pr_epsilonr}, and \ref{eq:pr_rhor}. The program runs four chains from different initial values, and checks for convergence by ensuring that the Gelman-Rubin statistic $\hat{R}$~\cite{gelman1992inference} is $<1.1$ and the effective sample size is $>100$; the number of MCMC iterations is increased until these convergence is achieved. The program \texttt{dms\_logoplot} in the \texttt{dms\_tools} package visualizes the posterior mean preferences via an extension to \texttt{weblogo}~\cite{crooks2004}.

>>>>

\subsubsection*{Implementation.}
The program \texttt{dms\_inferprefs} in the \texttt{dms\_tools} package infers the preferences by using \texttt{pystan}~\cite{pystan-software:2014} to perform MCMC over the posterior defined by the product of the likelihoods and priors in Equations \ref{eq:pr_nrerrpre}, \ref{eq:pr_nrerrpost}, \ref{eq:pr_nrpre}, \ref{eq:pr_nrpost}, \ref{eq:pr_pir}, \ref{eq:pr_mur}, \ref{eq:pr_epsilonr}, and \ref{eq:pr_rhor}. The program runs four chains from different initial values, and checks for convergence by ensuring that the Gelman-Rubin statistic $\hat{R}$~\cite{gelman1992inference} is $<1.1$ and the effective sample size is $>100$; the number of MCMC iterations is increased until these convergence is achieved. The program \texttt{dms\_logoplot} in the \texttt{dms\_tools} package visualizes the posterior mean preferences via an extension to \texttt{weblogo}~\cite{crooks2004}. The program \texttt{dms\_merge} can be used to average preferences inferred from different experimental replicates that have individually been analyzed by \texttt{dms\_inferprefs}, and the program \texttt{dms\_correlate} can be used to compute the correlations among inferences from different replicates.
